---
title: "Anomaly Detection"
author: "Brian Michira"
date: "9/10/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Research Question
To check whether the are anomalies in the given sales data.The objective of the task is fraud detection.

```{r}
# Load libraries
suppressWarnings(
        suppressMessages(if
                         (!require(tidyverse, quietly=TRUE))
                install.packages("tidyverse")))
library(tidyverse)
suppressWarnings(
        suppressMessages(if
                         (!require(anomalize, quietly=TRUE))
                install.packages("anomalize")))
library(anomalize)
suppressWarnings(
        suppressMessages(if
                         (!require(tibbletime, quietly=TRUE))
                install.packages("tibbletime")))
library(tibbletime)
suppressWarnings(
        suppressMessages(if
                         (!require(dplyr, quietly=TRUE))
                install.packages("dplyr")))
library(dplyr)
```

```{r}
#Load the data 
df <-read.csv("http://bit.ly/CarreFourSalesDataset")

```
```{r}
#checking the info
str(df)
```
```{r} 
#changing the data type
df$Date <- as.Date(df$Date, format ="%m/%d/%Y")
df$Date <- sort(df$Date, decreasing = FALSE)
df <- as_tbl_time(df, index = Date)
df <- df %>%as_period("daily")
head(df)
```
```{r}
#confirming if the data type has changed
str(df)
```

```{r}
#Check the shape
dim(df)
```
The data has 89 rows and 2 columns.

#Cleaning
```{r}
#checking for missing values
sum(is.na(df))
```
No missing values
```{r}
#checking for missing values
duplicated_rows<-df[duplicated(df),]
duplicated_rows
```
There are no duplicated rows.
```{r}
#checking for ouliers
boxplot(df$Sales)
```
There are no outliers

# Univariate Analysis

```{r}
# plot the histogram of Unit.price
ggplot(df, aes(x = Sales)) +
  geom_histogram(fill = "cyan", 
                 color = "black",
                 binwidth = 100) + 
  labs(title="Distribution of Sales",
       x = "Sales")
```
Distibution of sales are positively skewed showing that most of the values are greater than the mean.



```{r}
# Detecting our anomalies

df %>%
    time_decompose(Sales) %>%
    anomalize(remainder) %>%
    time_recompose() %>%
    plot_anomalies(time_recomposed = TRUE, ncol = 3, alpha_dots = 0.5)
```
There are no anomalies in our dataset.
