---
title: "Association Analysis"
author: "Brian Michira"
date: "9/10/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Research Question
To create association rules that will allow you to identify relationships between variables in the dataset.Being provided with a separate dataset that comprises groups of items that will be associated with others.
```{r}
# Loading the arules library
suppressWarnings(
        suppressMessages(if
                         (!require(arules, quietly=TRUE))
                install.packages("arules")))
library(arules)
```


```{r}
#Load the data and preview the head
path="http://bit.ly/SupermarketDatasetII"
supermarket <-read.transactions(path, sep = ",")
```

```{r}
# Verifying the object's class
class(supermarket)
```

```{r}
# Previewing our first 10 transactions
inspect(supermarket[1:10])
```
```{r}
# Creating a dataframe for the items
items<-as.data.frame(itemLabels(supermarket))
colnames(items) <- "Item"
head(items, 10) 
```
```{r}
# Generating a summary of the transaction dataset
summary(supermarket)
```
There are 7501 transactions in our dataset.The most purchased items were mineral water, eggs, spaghetti, french fries and chocolate. 

```{r}
# Exploring the frequency of some articles and checking the transaction percentages of the first 10 items.
itemFrequency(supermarket[, 1:10],type = "absolute")
round(itemFrequency(supermarket[, 1:10],type = "relative")*100,2)
```
```{r}
# plot the frequency of items
# Displaying top 10 most common items in the transactions dataset
itemFrequencyPlot(supermarket, topN = 10,col="cyan")
```
```{r}
# Frequency items whose relative importance is at least 10%
itemFrequencyPlot(supermarket, support = 0.1,col="purple")
```
```{r}
# Building a model based on association rules using the apriori function 
# We use Min Support as 0.001 and confidence as 0.8

rules <- apriori (supermarket, parameter = list(supp = 0.001, conf = 0.8))
rules
```
We found a set of 74 rules with a Min support of 0.001 and a  onfidence of 0.8.
```{r}
# However, in order to illustrate the sensitivity of the model to these two parameters, we will see what happens if we increase the support or lower the confidence level

# Building a apriori model with Min Support as 0.002 and confidence as 0.8.
rules2 <- apriori (supermarket,parameter = list(supp = 0.002, conf = 0.8)) 
rules2
```
Increasing the Min support we obtain a set of 2 rules.
```{r}
# Building apriori model with Min Support as 0.002 and confidence as 0.6.
rules3 <- apriori (supermarket, parameter = list(supp = 0.001, conf = 0.6)) 

rules3
```

Using a lower confidence of 0.6 we obtained 545 rules.

From the analysis we can see that using a higher level of support can make a mode lose interesting rules and reducing the confidence level increases the number of rules to quite an extent and many will not be useful.

```{r}
# Check the summaries of the rules
summary(rules)
```
```{r}
# Observing rules built in our model i.e. first 10 model rules
inspect(rules[1:10])
```
 If a customer buys frozen smoothie and spinach there is 88% confidence  of him or her purchasing mineral water or if one milk and pasta there is 85% confidence of him or her purchasing shrimp.
```{r} 
#plotting the top ten rules
library(arulesViz)
topRules <- rules[1:10]
plot(topRules)
```

```{r}
plot(topRules, method="graph")
```
The most porpular transaction was mushroom cream sauce and pasta.

```{r}
# Ordering these rules by a criteria such as the level of confidence then looking at the first five rules.

rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:5])
```
```{r}
#ordering by lift
rules<-sort(rules, by="lift", decreasing=TRUE)
inspect(rules[1:5])
```
```{r}
#ordering by support
rules<-sort(rules, by="support", decreasing=TRUE)
inspect(rules[1:5])
```



#Promotion
```{r}
# If we're interested in making a promotion relating to the sale of chocolate, 
# we could create a subset of rules concerning these products 

spaghetti <- subset(rules, subset = rhs %pin% "spaghetti")
 
# Then order by confidence
spaghetti<-sort(spaghetti, by="confidence", decreasing=TRUE)
inspect(spaghetti[1:5])
```
```{r}
# What if we wanted to determine items that customers might buy who have previously bought spaghetti?
# Subset the rules
spaghetti <- subset(rules, subset = lhs %pin% "spaghetti")

# Order by confidence
spaghetti<-sort(spaghetti, by="confidence", decreasing=TRUE)

# inspect top 5
inspect(spaghetti[1:5])
```
